
# World {#world}


## Daily Incidence, Recovery, Deaths {#incidence-recovery-death}

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
# source("global_incidence_recovery_deaths.R")
source("covid19_global.R")
source("load_functions.R")
```

__Report last updated:__ `r Sys.Date()`

__Data last refreshed:__ : `r data_last_refreshed`

```{r global-incidence-recovery-deaths, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}

p <- global %>%
  group_by(date) %>%
  summarise(
    daily_cases = sum(daily_cases)
    , daily_recov = sum(daily_recov)
    , daily_deaths = sum(daily_deaths)
    , daily_recov_death_ratio = (daily_recov/daily_deaths)
  ) %>%
  gather(metric, n, daily_recov, daily_cases, daily_deaths, factor_key = TRUE) %>%
  mutate(
    metric = factor(metric, levels = c("daily_deaths", "daily_recov", "daily_cases"))
  ) %>%
  ggplot(aes(x=date, y=n, color=metric))+
  # geom_point(aes(y=rollmean(n, 7, na.pad=TRUE))) +
  geom_point() +
  geom_smooth() +
  theme_minimal(base_size = 14) +
  theme(legend.position="top") +
  ggtitle("GLOBAL daily incidence, recovery, deaths", data_last_refreshed ) +
  xlab("Date") +
  ylab("Daily cases")+
  labs(
    title= "GLOBAL daily incidence, recovery, deaths  (log scale)",
    caption = "SOURCE: Johns Hopkins Coronavirus Resource Center"
    )
p = p + scale_y_continuous(labels=comma_format(), trans='log10') +
  scale_color_manual(name="Metric", 
                   breaks=c("daily_deaths", 
                                  "daily_recov", "daily_cases"),
                   labels=c("Daily deaths", 
                                  "Daily recovery",
                                  "Daily cases"),
                   values=ggplotColours(n = 3)) +
   guides(fill=guides(colour=guide_legend(ncol=3,nrow=1,byrow=TRUE)))

suppressMessages(print(p))

```


## Case Fatality Rate {#world-case-fatality}

Case fatality rate is calculated by dividing the number of deaths from a specified disease over a defined period of time by the number of individuals diagnosed with the disease during that time; the resulting ratio is then multiplied by 100 to yield a percentage [@cfr_britanica].

$$CFR = \frac{\# Deaths}{\#Infected} \times 100 \,\, per \,\,unit \,\, time$$

```{r, global-case-fatality-rate, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}

df <- case_fatality(global)

p <- df %>%
  ggplot(aes(x=date, y=n, color=metric))+
  geom_point() +
  geom_smooth() +
  theme_minimal(base_size = 14) +
  theme(legend.position="top") +
  ggtitle("GLOBAL case fatality rate", data_last_refreshed ) +
  xlab("Date") +
  ylab("Case fatality rate (%)")+
  labs(
    title= "GLOBAL case fatality rate",
    caption = "SOURCE: Johns Hopkins Coronavirus Resource Center"
  )
p = p + scale_y_continuous(labels=comma, expand=c(0.5, 0.5)) +
  scale_color_discrete(name="Metric",
                       breaks=c("cum_case_fatality_rate",
                                "daily_case_fatality_rate"),
                       labels=c("Cumulative case fatality",
                                "Daily case fatality")) +
  guides(fill=guides(colour=guide_legend(ncol=3,nrow=1,byrow=TRUE)))

suppressMessages(print(p))

```

### Case Fatality Rate (select countries)


```{r, case-fatality-rate-selected, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}

country_list = c("Bangladesh", "US", "United Kingdom", "Canada", "France", 
                 "Italy", "India", "Pakistan", "")
 
df <- global %>%
  filter(country %in% country_list, cum_cases >0) %>%
  group_by(country, date) %>%
    summarise(
      cum_case_fatality_rate = (sum(cum_deaths)/sum(cum_cases)) * 100
      , daily_case_fatality_rate = (sum(daily_deaths)/sum(cum_cases)) * 100
    ) %>%
    group_by(country) %>%
  mutate(
   day = row_number()
  ) %>%
  gather(metric, n, cum_case_fatality_rate, daily_case_fatality_rate, factor_key = TRUE)


p <- df %>%
  ggplot(aes(x=day, y=n, color=metric))+
  geom_point() +
  geom_smooth() +
  theme_minimal(base_size = 14) +
  theme(legend.position="top") +
  ggtitle("GLOBAL case fatality rate", data_last_refreshed ) +
  xlab("Day since first case") +
  ylab("Case fatality rate (%)")+
  labs(
    title= "CASE fatality rate (selected countries)",
    caption = "SOURCE: Johns Hopkins Coronavirus Resource Center"
  )
p = p + scale_y_continuous(labels=comma, expand=c(0.5, 0.5)) +
  scale_color_discrete(name="Metric",
                       breaks=c("cum_case_fatality_rate",
                                "daily_case_fatality_rate"),
                       labels=c("Cumulative case fatality",
                                "Daily case fatality")) +
  guides(fill=guides(colour=guide_legend(ncol=3,nrow=1,byrow=TRUE))) +
  facet_wrap( ~ country, ncol = 2)

suppressMessages(print(p))


```

## Confirmed Cases 


```{r, global-confirmed, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}
# Cumulative Cases since first identified case

# Confirmed for Bangladesh
df_bd <- global %>%
  filter(
    country %in% c('Bangladesh')
  ) %>%
  select(country, date, cum_cases) %>%
  arrange(
    country, date
  ) 

# Top 10 countries
n_world <- global %>%
  filter(country != 'Bangladesh(unoff)' ) %>%
  group_by(country) %>%
  summarise(
    max_n_cases = max(cum_cases)
  ) %>%
  arrange(desc(max_n_cases)) %>%
  top_n(6, max_n_cases)

# Top 10 countries
n_bd <- df_bd %>%
  group_by(country) %>%
  summarise(
    max_n_cases = max(cum_cases)
  )


# Combined n countries
n <- rbind(n_world, n_bd)

# Reset the origin to the first reported cases
case_summary <- global %>%
  filter(country %in% n$country, cum_cases >0) %>%
  arrange(country, date, cum_cases) %>%
  group_by(country) %>%
  mutate(
   day = row_number()
  )
data_last_refreshed = max(case_summary$date)

p <- ggplot(case_summary, aes(x = day, y = cum_cases, color=country)) +
  geom_line(size=1.1) +
  geom_point(data = case_summary[case_summary$country == 'Bangladesh' & ymd(case_summary$date)==data_last_refreshed, ],
             aes(x=day, y=cum_cases), color='red', size=5) +
  geom_text_repel(size=5, aes(label=comma_format()(cum_cases)),
                  function(case_summary) case_summary[ymd(case_summary$date) == data_last_refreshed, ])+
  theme_minimal(base_size = 14) +
  theme(legend.position="top") +
  ggtitle("COVID19 cumulative cases since first case \n(top 6 countries)", data_last_refreshed ) +
  xlab("Days since first confirmed case") +
  ylab("Cumulative cases")+
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center.")

p + scale_color_discrete(name="Country") +
  scale_y_continuous(labels=comma, trans = 'log10')

# +
#   guides(fill=guides(colour=guide_legend(ncol=1,nrow=1,byrow=TRUE))) +
#   scale_y_continuous(labels = function(x) format(x, scientific = FALSE))



```


```{r, echo=FALSE}

# Summary of cases till date by country
world_summary <- global %>%
  # filter(country != 'Bangladesh(unoff)') %>%
  filter(ymd(date) == data_last_refreshed) %>%
  group_by(country) %>%
  summarise(
    Confirmed = max(cum_cases)
  ) %>%
  top_n(10, Confirmed) %>%
  arrange(desc(Confirmed))

confirmed_bd <- bd_unoff %>%
  filter(ymd(date) == data_last_refreshed) %>%
  group_by(country) %>%
  summarise(
    Confirmed = max(cum_cases)
  ) %>%
  arrange(desc(Confirmed))

# Confirmed (world + Bangladesh)

world_plus_bd <- rbind(world_summary, confirmed_bd)

world_plus_bd %>%
  kable(format = "html", caption=paste("Top 10 countries by confirmed cases as of", data_last_refreshed)) %>%
  kable_styling(position = "center")


```


## Deaths since first death


```{r, global-deaths-since-first-death, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE}
# Cumulative Cases since first identified case

b <- bd_unoff
a <- global %>% filter(country %in% "Bangladesh", cum_deaths > 0) %>%
  select(-c(daily_cases, daily_deaths, daily_recov))

bd <- bind_rows(a,b) %>%
  arrange(country, date) %>%
  mutate(
   day = row_number()
  )

# Top 10 countries
n_world <- global %>%
  group_by(country) %>%
  summarise(
    max_n_cases = max(cum_deaths)
  ) %>%
  arrange(desc(max_n_cases)) %>%
  top_n(6, max_n_cases)


df_world <- global %>%
  filter(country %in% n_world$country, cum_deaths >0) %>%
  select(date, country, cum_cases, cum_deaths, cum_recov)

df_world_bd <- bind_rows(bd, df_world)


# Reset the origin to the first reported cases
case_summary <- df_world_bd %>%
  filter(cum_deaths >0) %>%
  arrange(country, date, cum_deaths) %>%
  group_by(country) %>%
  mutate(
   day = row_number()
  )


max_date_bd_unoff <- case_summary %>%
  filter(country == "Bangladesh(unoff)") %>% group_by(country) %>%
  summarise(date=max(date))

max_date_bd_off <- case_summary %>%
  filter(country == "Bangladesh") %>% group_by(country) %>%
  summarise(date=max(date))


p <- ggplot(case_summary, aes(x = day, y = cum_deaths, color=country)) +
  geom_line(size=1.1) +
  geom_point(data = case_summary[case_summary$country == 'Bangladesh(unoff)' & case_summary$date==max_date_bd_unoff$date, ], aes(x=day, y=cum_deaths), color='red', size=5) +
  geom_point(data = case_summary[case_summary$country == 'Bangladesh' & case_summary$date==max_date_bd_off$date, ], aes(x=day, y=cum_deaths), color='red', size=5) +
  geom_text_repel(size=5, aes(label=country),
                  function(case_summary) case_summary[case_summary$date == max_date_bd_off$date, ])+
  
  theme_minimal(base_size = 14) +
  theme(legend.position="top") +
  ggtitle("COVID19 cumulative deaths since first case", data_last_refreshed ) +
  xlab("Days since first case") +
  ylab("Cumulative deaths")+
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center.")

p + scale_color_discrete(name="Country") +
  scale_y_continuous(labels=comma_format(), trans = 'log10')

  # +
#   guides(fill=guides(colour=guide_legend(ncol=1,nrow=1,byrow=TRUE))) +
#   scale_y_continuous(labels = function(x) format(x, scientific = FALSE))



```

## Deaths since first confirmed case



```{r, global-deaths-since-first-case, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE}
# Cumulative Cases since first identified case

b <- bd_unoff
a <- global %>% filter(country %in% "Bangladesh", cum_cases > 0) %>%
  select(-c(daily_cases, daily_deaths, daily_recov))

bd <- bind_rows(a,b) %>%
  arrange(country, date) %>%
  mutate(
   day = row_number()
  )

# Top 10 countries
n_world <- global %>%
  group_by(country) %>%
  summarise(
    max_n_cases = max(cum_deaths)
  ) %>%
  arrange(desc(max_n_cases)) %>%
  top_n(6, max_n_cases)


df_world <- global %>%
  filter(country %in% n_world$country, cum_cases >0) %>%
  select(date, country, cum_cases, cum_deaths, cum_recov)

df_world_bd <- bind_rows(bd, df_world)


# Reset the origin to the first reported cases
case_summary <- df_world_bd %>%
  filter(cum_cases >0) %>%
  arrange(country, date, cum_deaths) %>%
  group_by(country) %>%
  mutate(
   day = row_number()
  )


max_date_bd_unoff <- case_summary %>%
  filter(country == "Bangladesh(unoff)") %>% group_by(country) %>%
  summarise(date=max(date))

max_date_bd_off <- case_summary %>%
  filter(country == "Bangladesh") %>% group_by(country) %>%
  summarise(date=max(date))


p <- ggplot(case_summary, aes(x = day, y = cum_deaths, color=country)) +
  geom_line(size=1.1) +
  geom_point(data = case_summary[case_summary$country == 'Bangladesh(unoff)' & case_summary$date==max_date_bd_unoff$date, ], aes(x=day, y=cum_deaths), color='red', size=5) +
  geom_point(data = case_summary[case_summary$country == 'Bangladesh' & case_summary$date==max_date_bd_off$date, ], aes(x=day, y=cum_deaths), color='red', size=5) +
  geom_text_repel(size=5, aes(label=country),
                  function(case_summary) case_summary[case_summary$date == max_date_bd_off$date, ])+
  
  theme_minimal(base_size = 14) +
  theme_minimal(base_family="Siyam Rupali") +
  theme(legend.position="none") +
  ggtitle("COVID19 cumulative deaths since first confirmed case", data_last_refreshed ) +
  xlab("Days since first confirmed case") +
  ylab("Cumulative deaths")+
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center and other.")

p + scale_color_discrete(name="Country") +
  scale_y_continuous(labels=comma_format(), trans = 'log10') +
  scale_x_continuous(limits=c(0,max(case_summary$day)+5), breaks=seq(0,max(case_summary$day)+5,10))

  # +
#   guides(fill=guides(colour=guide_legend(ncol=1,nrow=1,byrow=TRUE))) +
#   scale_y_continuous(labels = function(x) format(x, scientific = FALSE))



```




```{r, echo=FALSE}

# Summary of cases till date by country
world_summary <- global %>%
  filter(ymd(date) == data_last_refreshed) %>%
  group_by(country) %>%
  summarise(
    Confirmed = max(cum_deaths)
  ) %>%
  top_n(10, Confirmed) %>%
  arrange(desc(Confirmed))

confirmed_bd <- bd_unoff %>%
  filter(ymd(date) == max_date_bd_unoff$date) %>%
  group_by(country) %>%
  summarise(
    Confirmed = max(cum_deaths)
  ) %>%
  arrange(desc(Confirmed))

# Confirmed (world + Bangladesh)

world_plus_bd <- rbind(world_summary, confirmed_bd)

world_plus_bd %>%
  kable(format = "html", caption=paste("Top 10 countries by confirmed deaths as of", data_last_refreshed)) %>%
  kable_styling(position = "center")


```


### Cases and deaths compared on a specific day

```{r, echo=FALSE}
bd_day_indicator = df_world_bd %>% filter(country == 'Bangladesh', 
                                                date == max(date)) %>%
  pull(day)
```
Bangladesh has entered on day `r bd_day_indicator` since first confirmed case.


```{r, echo=FALSE}

# Summary of deaths till date by country on the specific day

bd_world_summary <- df_world_bd %>%
  group_by(country) %>%
  mutate(
   day = row_number()
  ) %>%
  filter(day == bd_day_indicator) %>%
  group_by(country) %>%
  summarise(
    `Total cases` = max(cum_cases),
    `Total deaths` = max(cum_deaths),
    Country = country
  ) %>%
  select(Country, `Total cases`, `Total deaths`) %>%
  arrange(desc(`Total deaths`))

bd_world_summary %>%
  kable(format = "html", caption=paste0("Bangladesh in the world: Total cases and deaths compared on day ", bd_day_indicator)) %>%
  kable_styling(position = "center") %>%
  row_spec(which(bd_world_summary$Country %in% c('Bangladesh', 'Bangladesh(unoff)')), bold = T, color = "white", background = "#f49a86")

```

## Recovery-to-death ratio {#recovery-death-ratio}


The recovery to deaths ratio (RDR) is calculated as

$$
RDR = \frac{\# Cumulative \,\, daily \,\, recovered}{\# Cumulative \,\, daily \,\, deaths}
$$


__How to interpret__: The ratio of cumulative number of patients recovered divided by the cumulative number of deaths over time can tell the proportion of people recovering compared to dying. An upward trend may indicate more recovery than deaths. 


```{r, recovery-to-death-ratio-by-country, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}

# Global top 5 countries by deaths --------------

# Top n countries
n_world <- global %>%
  group_by(country) %>%
  summarise(
    deaths = max(cum_deaths)
    , recovered = max(cum_recov)
  ) %>%
  arrange(country) %>%
  top_n(5, deaths)

# Combined n countries
n <- n_world

df <- global %>%
  filter(country != "Bangladesh(unoff)") %>%
  mutate(
    # cum_recov_death_ratio = cum_recov/cum_deaths
    cum_recov_death_ratio = cum_recov/cum_deaths
  ) %>%
  filter(is.finite(cum_recov_death_ratio)) %>%
  filter(country %in% n$country, cum_recov_death_ratio >= 0) %>%
  arrange(country, date, cum_recov_death_ratio) %>%
  group_by(country) %>%
  mutate(
    day = row_number()
  )
   p <- ggplot(df, aes(x = day, y = cum_recov_death_ratio, color=country)) +
      geom_point(size=1) + geom_smooth() +
     geom_text_repel(size=5, aes(label=country),
                     function(df) df[ymd(df$date) == data_last_refreshed, ])+
     theme_minimal(base_size = 14)

  p <-  p + ggtitle("Recovery-to-death ratio \n(Top 5 countries by death, excluding China)", data_last_refreshed ) +
      xlab("Days since first death") +
      ylab("Cumulative recovery / cumulative deaths") +
      theme(legend.position="top") +
      labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center") +
      scale_color_discrete(name="Country")
  
suppressMessages(print(p))


```


### Recovery-to-death ratio (South and South-East Asia)


```{r, recovery-to-death-ratio-south-asia, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}

country_list = c("Bangladesh", "India", "Singapore", "Nepal", "Pakistan",
                   "Bhutan", "Sri Lanka", "Malaysia", "Maldives", 
                   "Indonesia", "Thailand")

# Top n countries
n_world <- global %>%
  filter(country %in% country_list) %>%
  group_by(country) %>%
  summarise(
    deaths = max(cum_deaths)
    , recovered = max(cum_recov)
  ) %>%
  arrange(country) %>%
  top_n(10, deaths)

# Combined n countries
n <- n_world

df <- global %>%
  mutate(
    cum_recov_death_ratio = cum_recov/cum_deaths
  ) %>%
  filter(country %in% country_list) %>%
  filter(is.finite(cum_recov_death_ratio)) %>%
  filter(country %in% n$country, cum_recov_death_ratio >= 0) %>%
  arrange(country, date, cum_recov_death_ratio) %>%
  group_by(country) %>%
  mutate(
    day = row_number()
  )

data_last_refreshed = max(df$date)
# plot
p <- ggplot(df, aes(x = day, y = cum_recov_death_ratio, color=country)) +
  geom_point(size=1) + geom_smooth() +
  geom_text_repel(size=5, aes(label=country),
                  function(df) df[ymd(df$date) == data_last_refreshed, ])+
  theme_minimal(base_size = 14)

p <-  p + ggtitle("Recovery-to-death ratio \n(South Asian countries)", data_last_refreshed ) +
  xlab("Days since first death") +
  ylab("Cumulative recovery / cumulative deaths") +
  theme(legend.position="top") +
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center") +
  scale_color_discrete(name="Country")

suppressMessages(print(p))

```

### Recovery-to-death ratio (Middle East)


```{r, recovery-to-death-ratio-middle-east, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}

country_list = c("Saudi Arabia", "Iraq", "Iran", "Israel",  "United Arab Emirates")

# Top n countries
n_world <- global %>%
  filter(country %in% country_list) %>%
  group_by(country) %>%
  summarise(
    deaths = max(cum_deaths)
    , recovered = max(cum_recov)
  ) %>%
  arrange(country) %>%
  top_n(10, deaths)

# Combined n countries
n <- n_world

df <- global %>%
  mutate(
    cum_recov_death_ratio = cum_recov/cum_deaths
  ) %>%
  filter(country %in% country_list) %>%
  filter(is.finite(cum_recov_death_ratio)) %>%
  filter(country %in% n$country, cum_recov_death_ratio >= 0) %>%
  arrange(country, date, cum_recov_death_ratio) %>%
  group_by(country) %>%
  mutate(
    day = row_number()
  )

# plot
p <- ggplot(df, aes(x = day, y = cum_recov_death_ratio, color=country)) +
  geom_point(size=1) + geom_smooth() +
  geom_text_repel(size=5, aes(label=country),
                  function(df) df[ymd(df$date) == data_last_refreshed, ])+
  theme_minimal(base_size = 14)

p <-  p + ggtitle("Recovery-to-death ratio \n(Middle Eastern countries)", data_last_refreshed ) +
  xlab("Days since first death") +
  ylab("Cumulative recovery / cumulative deaths") +
  theme(legend.position="top") +
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center") +
  scale_color_discrete(name="Country")

suppressMessages(print(p))

```


## 3-day % changes {#perc-change-top6}



```{r, 3day-perc-change, fig.width=8, fig.height=9, echo = FALSE, warning=FALSE}

# Global top 6 countries by deaths --------------

# Top 6 countries by deaths
global_top_deaths <- global %>%
  filter(!country %in% c('China')) %>%
  group_by(country) %>%
  summarise(
    max_n_cases = max(cum_deaths)
  ) %>%
  arrange(desc(max_n_cases)) %>%
  top_n(6, max_n_cases)

p <- global_gr_lagx(global, global_top_deaths$country, 3)
p <- p + theme(legend.position="none") +
  ggtitle("COVID19 3-day % change in select metrics \n(Top 6 countries by death, excluding China)", data_last_refreshed ) +
  xlab("Days since first death") +
  ylab("% change") +
  theme(legend.position="bottom") +
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center")

p = p + scale_color_discrete(name="Metric",
                         breaks=c("case_gr_rate", "death_gr_rate"),
                         labels=c("New cases", "Deaths")) +
  guides(fill=guides(colour=guide_legend(ncol=2,nrow=1,byrow=TRUE))) +
  facet_wrap( ~ country, ncol = 2) +
  ylim(c(-20, 300))
p

```


The percent change is calculated as 

$$
PC = \frac{New - Original}{Original} * 100
$$
where `New` is the number of cases on a given date, and `Original` is the number 3 days prior to that. 

__How to interpret__: The 3-day percent change tells how much does a metric have changed (%) compared to what it was 3-days ago. A downward trending line suggests a slow growth.

