# Bangladesh {#bangladesh}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# source("covid19_bd.R")
source("covid19_global.R")
library(incidence)
```

This report is created based on a publicly available data. The data are available on a Google Sheet. Please see the Data Source section for link. 
Please note that publicly available data are __not official__ and __MAY BE UNRELIABLE__. I personally did not verify them. I am using them for educational purposes. Use at your discretion.

## Epidemic curve and search for peak

The epidemic curve for Bangladesh and a tentative peak. The red line in the plot indicates a potential peak based on available data. IF the peak (red line) is at the far right, then we may not have reached a peak yet. We must have sustained decrease of incidences to be sure about any peak. 


```{r, covid19-bangladesh-epidemic-curve-peak, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}
# BD unofficial cases
date_cut_off = ymd("2020-05-10")
date_data_refreshed = max(bd_unoff$date)

bd_unoff_cases <- bd_unoff %>%
  mutate(
    country = "Bangladesh(unoff)"
    , date = as.Date(date, "%d/%m/%Y")
    , daily_cases = c(0, diff(cum_cases))
  ) %>%
  # filter(date <= date_cut_off) %>%
  select(date, country, cum_cases, daily_cases) %>%
  drop_na()

bd_incidence_object <- bd_unoff_cases %>% uncount(daily_cases)
# Crate incidence object using incidence package
bd_incidence_object <- incidence::incidence(bd_incidence_object$date)
# Finding peak
bd_incidence_peak <- incidence::find_peak(bd_incidence_object)
plot(bd_incidence_object) + 
  geom_vline(xintercept = bd_incidence_peak, col = "red", 
             lty = 2, size=1) +  
  labs(title = paste0("Daily unofficial new infections in Bangladesh as of ", max(bd_unoff_cases$date)), 
       subtitle = "(Red dashed line indicates date of peak infections)") + 
  #theme(text = element_text(size = 12))
  theme(text = element_text(size = 12), axis.text.x = element_text(angle = 90, hjust = 1))

```


We fit a log linear model to estimate the doubling time (growh phase) and halving time (decline phase, when that happens). The estimated fitted line (solid line) with 95\% confidence interval (dotten lines) are superimposed on the epidemic curve below.


```{r, covid19-bangladesh-epidemic-curve-fitted, fig.width=8, fig.height=6, echo = FALSE, warning=FALSE}
# Fitting the log-linear models to the growth section and the decay section
# bd_incidence_fit <- incidence::fit(bd_incidence_object, split = bd_incidence_peak)
bd_incidence_fit <- incidence::fit(bd_incidence_object)
# plot the incidence data and the model fit
plot(bd_incidence_object) %>% add_incidence_fit(bd_incidence_fit) + 
  labs(title = "Observed and fitted incidence of COVID-19", 
       subtitle = "Bangladesh, 2020") + 
  theme(text = element_text(size = 12))
```

The current doubling time as of `r today()` is `r round(get_info(bd_incidence_fit, "doubling"), 1)` days with a 95% confidence interval of (`r round(get_info(bd_incidence_fit, "doubling.conf"), 1)`) days. 

### Estimating the Overall Reproduction Number 

The log-linear model estimates the $R_0$ as follows:

```{r, echo=FALSE}

# Getting the R0 from log-linear model ----

library(distcrete)
library(epitrix)

mu <- 7.5  # days
sigma <- 3.4  # days
param <- gamma_mucv2shapescale(mu, sigma/mu)

w <- distcrete("gamma", interval = 1, shape = param$shape, scale = param$scale, w = 0)

# growth_R0 <- lm2R0_sample(bd_incidence_fit$before$model, w)
growth_R0 <- lm2R0_sample(bd_incidence_fit$model, w)
```

```{r, echo=FALSE}

hist(growth_R0, col = "grey", border = "white", main = "Distribution of R0")
summary(growth_R0)

```


### Estimating the Instantaneous Reproduction Number ($R_e$)

The estimation of effective reproduction number involves the distribution of serial interval (SI). Based on the literature (REF to be added), a gamma distribution with mean 7.5 and standard deviation 3.5

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(EpiEstim)

# custom results plotting function to avoid the ugly
# TableGrob messages returned by the plotting function in the
# EpiEstim package
plot_Ri <- function(estimate_R_obj) {
  p_I <- plot(estimate_R_obj, "incid", add_imported_cases = TRUE)  # plots the incidence
  p_SI <- plot(estimate_R_obj, "SI")  # plots the serial interval distribution
  p_Ri <- plot(estimate_R_obj, "R")
  return(gridExtra::grid.arrange(p_I, p_SI, p_Ri, ncol = 2))
}

bd_confirmed_cases <- bd_unoff_cases %>%
  mutate(
    daily_cases = c(0, diff(cum_cases))
  )
  
  
bd_res_parametric_si <- 
  estimate_R(bd_confirmed_cases$daily_cases, 
             method = "parametric_si", 
             config = make_config(list(mean_si = 7.5, std_si = 3.5))
)

plot_Ri(bd_res_parametric_si)
```


### Simulating the $R_e$

Now we perform a simulation of different SI distribution to obtain the $R_e$. For this, we vary the parameters of gamma distribution and obtain the following.


```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Recalculating with varying SI distribution parameters 
# we get these results:

config = make_config(list(mean_si = 7.5, std_mean_si = 3, 
                                       min_mean_si = 2, max_mean_si =10.0,
                                       std_si = 3.5, std_std_si = 1,
                                       min_std_si = 0.5, max_std_si = 4))
bd_res_uncertain_si <- 
  estimate_R(bd_confirmed_cases$daily_cases, method = "uncertain_si", 
             config = config)

plot_Ri(bd_res_uncertain_si)

```

```{r, echo=FALSE}
# print the last 7 days of data
bd_res_uncertain_si$R %>%
  tail(7) %>%
  select(t_start, t_end, `Mean(R)`, `Std(R)`,  `Quantile.0.025(R)`, `Quantile.0.975(R)`) %>%
  kable(format = "html", caption=paste("Instantaneous Effective Reproduction Number by Week as of ", date_data_refreshed)) %>%
  kable_styling(position = "center")

R_CI = tail(bd_res_uncertain_si$R, 1)
R_mean = round(as.numeric(R_CI["Mean(R)"]), 2)
R_LCI = round(as.numeric(R_CI["Quantile.0.025(R)"]), 2)
R_UCI = round(as.numeric(R_CI["Quantile.0.975(R)"]), 2)
```

The effective reproduction number based on the last week's incidences show an average of `r R_mean` with a 95\% confidence interval (`r R_LCI`, `r R_UCI`).

### Cumulative Deaths (official and unofficial)

```{r, covid19-bangladesh-deaths, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE}
b <- bd_unoff
a <- global %>% filter(country %in% "Bangladesh") %>%
  select(-c(daily_cases, daily_deaths, daily_recov), cum_cases)
bd <- bind_rows(a,b) %>%
  filter(cum_cases >0 ) %>%
  arrange(country, date) %>%
  mutate(
   day = row_number()
  )
  
# Plot
p <- ggplot(bd , aes(x = day, y = cum_deaths, color=country)) +
  geom_line(size=1.1) +
  # geom_text_repel(size=5, aes(label=country),
  #                 function(a) a[ymd(a$date) == max(a$date), ])+
  theme_minimal(base_size = 14) +
  theme(legend.position="top") +
  ggtitle("Bangladesh: COVID19 cumulative deaths",
          data_last_refreshed ) +
  xlab("Day since first confirmed case") +
  ylab("Cumulative deaths") +
  theme(legend.position="top") +
  labs(caption = "SOURCE: Bangladesh(unofficial) data compiled from newspaper reports")
p + scale_y_continuous(labels=comma, trans = "log10") 
  # guides(fill=guides(colour=guide_legend(ncol=2,nrow=1,byrow=TRUE)))
```


```{r, echo=FALSE}
bdsummary <- bd %>%
  filter(date == max(date)) %>% 
  group_by(country) %>%
  summarise(
    `Total Deaths` = max(cum_deaths)
  ) %>%
  arrange(desc(`Total Deaths`))
  
bdsummary %>%
  kable(format = "html", caption=paste("Total number of deaths as of ", max(bd$date))) %>%
  kable_styling(position = "center")
```

## Bangladesh in South Asia

### Infection

```{r, incidences-bd-south-asia, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE, message=FALSE}
# source("covid19_deaths.R")
# 
# bd_cases <- bd_unoff %>% filter(country == "Bangladesh(unoff)")
#   select(country, date, cum_cases)
df_south_asia <- global %>%
  filter(
    country %in% c("Bangladesh", "India", "Singapore", "Nepal", "Pakistan",
                   "Bhutan", "Sri Lanka", "Malaysia",
                   "Indonesia")
  ) %>%
  select(country, date, cum_cases) %>%
  group_by(country, date) %>%
  summarize(
    cum_cases = sum(cum_cases)
  )
# Join with BD data
# df_south_asia <- bind_rows(df_south_asia, bd_cases)
n_south_asia <- df_south_asia %>%
  group_by(country) %>%
  summarise(
    max_n = max(cum_cases)
  ) %>%
  arrange(desc(max_n)) %>%
  top_n(10, max_n)
# Combined n countries
# n_south_ais <- rbind(n_world, n_bd)
cases_south_asia <- df_south_asia %>%
  filter(country %in% n_south_asia$country, cum_cases >0) %>%
  arrange(country, cum_cases) %>%
  group_by(country) %>%
  mutate(
    day = row_number()
  )
```

```{r, covid19-cases-south-asia, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE, message=FALSE}
date_data_refreshed = max(cases_south_asia$date)
# Plot
ggplot(cases_south_asia, aes(x = day, y = cum_cases, color=country)) +
  geom_line(size=1.1) +
  geom_point(data = cases_south_asia[cases_south_asia$country == 'Bangladesh' & cases_south_asia$date==max(cases_south_asia$date), ], aes(x=day, y=cum_cases), color='red', size=5) +
  geom_text_repel(size=5, aes(label=country),
                  function(cases_south_asia) cases_south_asia[cases_south_asia$date == max(cases_south_asia$date), ])+
  theme_minimal(base_size = 14) +
  theme(legend.position = "none") +
  # theme_void() + 
  theme(legend.position="none") +
  ggtitle( paste("COVID19 cumulative incidences since first incidence \n(Bangladesh's peers)\n", date_data_refreshed)) +
  xlab("Days since first incidence") +
  ylab("Cumulative cases") +
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center.")+
  scale_y_continuous(labels=comma, trans = "log10")
```


```{r, echo=FALSE}
# Summary of Infections till date by country
d <- bd %>%
  select(country, date, cum_cases)
df_south_asia <- global %>%
  filter(
    country %in% c("India", "Singapore", "Nepal", "Pakistan",
                   "Bhutan", "Sri Lanka", "Malaysia", 
                   "Indonesia")
  ) %>%
  select(country, date, cum_cases) %>%
  group_by(country, date) %>%
  summarize(
    cum_cases = sum(cum_cases)
  )
df_south_asia <- bind_rows(df_south_asia, d)
n_south_asia <- df_south_asia %>%
  group_by(country) %>%
  summarise(
    max_n = max(cum_cases)
  ) %>%
  arrange(desc(max_n)) %>%
  top_n(20, max_n)
sa_summary_cases <- df_south_asia %>%
  group_by(country) %>%
  summarise(
    `Total Cases` = max(cum_cases)
  ) %>%
  arrange(desc(`Total Cases`))
sa_summary_cases %>%
  kable(format = "html", caption=paste("Total number of cases as of ", date_data_refreshed)) %>%
  kable_styling(position = "center")
```

### Deaths

```{r, southasia-deaths, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE, message=FALSE}
d <- bd %>%
  select(country, date, cum_cases, cum_deaths)
df_south_asia <- global %>%
  select(country, date, cum_deaths, cum_cases) %>%
  filter(
    country %in% c("India", "Singapore", "Nepal", "Pakistan",
                   "Bhutan", "Sri Lanka", "Malaysia", 
                   "Indonesia"),
    cum_cases > 0
  ) %>%
  group_by(country, date) %>%
  summarize(
    cum_deaths = sum(cum_deaths),
    cum_cases = sum(cum_cases)
  )
df_south_asia <- bind_rows(df_south_asia, d)
n_south_asia <- df_south_asia %>%
  group_by(country) %>%
  summarise(
    max_n = max(cum_deaths)
  ) %>%
  arrange(desc(max_n)) %>%
  top_n(10, max_n)
# Combined n countries
# n_south_ais <- rbind(n_world, n_bd)
deaths_south_asia <- df_south_asia %>%
  filter(country %in% n_south_asia$country) %>%
  arrange(country, cum_deaths) %>%
  group_by(country) %>%
  mutate(
    day = row_number()
  )
```


```{r, covid19-death-south-asia, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE, message=FALSE}
max_date_bd_unoff <- deaths_south_asia %>%
  filter(country == "Bangladesh(unoff)") %>% group_by(country) %>%
  summarise(date=max(date))
max_date_bd_off <- deaths_south_asia %>%
  filter(country == "Bangladesh") %>% group_by(country) %>%
  summarise(date=max(date))
# Plot
ggplot(deaths_south_asia, aes(x = day, y = cum_deaths, color=country)) +
  geom_line(size=1.1) +
  geom_point(data = deaths_south_asia[deaths_south_asia$country == 'Bangladesh' & deaths_south_asia$date==max_date_bd_off$date, ], aes(x=day, y=cum_deaths), color='red', size=5) +
  geom_point(data = deaths_south_asia[deaths_south_asia$country == 'Bangladesh(unoff)' & deaths_south_asia$date==max_date_bd_unoff$date, ], aes(x=day, y=cum_deaths), color='red', size=5) +
  geom_text_repel(size=5, aes(label=country),
                  function(deaths_south_asia) deaths_south_asia[deaths_south_asia$date == max_date_bd_off$date, ])+
  # theme_minimal(base_size = 14, base_family="Siyam Rupali") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none") +
  # theme_void() + 
  theme(legend.position="none") +
  ggtitle( paste("COVID19 cumulative deaths since first incidence \n(Bangladesh's peers)", date_data_refreshed)) +
  xlab("Days since first incidence") +
  ylab("Cumulative deaths") +
  labs(caption = "SOURCE: Johns Hopkins Coronavirus Resource Center. Bangladesh(unoff) data compiled from news reports")+
  scale_y_continuous(labels=comma, trans = "log10")
```


```{r, echo=FALSE}
# Summary of deaths till date by country
sa_summary <- deaths_south_asia %>%
  filter(date == max(date)) %>%
  group_by(country) %>%
  summarise(
    `Total Deaths` = max(cum_deaths)
  ) %>%
  arrange(desc(`Total Deaths`))
sa_summary %>%
  kable(format = "html", caption=paste("Total number of deaths as of ", date_data_refreshed)) %>%
  kable_styling(position = "center")
```


### Cases and deaths compared on a specific day

```{r, echo=FALSE}
bd_day_indicator = df_world_bd %>% filter(country == 'Bangladesh', 
                                                date == max(date)) %>%
  pull(day)
```
Bangladesh has entered into day `r bd_day_indicator` since first confirmed case.


```{r, echo=FALSE}
# Summary of deaths till date by country on the specific day
bd_sa_summary <- df_south_asia %>%
  group_by(country) %>%
  mutate(
   day = row_number()
  ) %>%
  filter(day == bd_day_indicator) %>%
  group_by(country) %>%
  summarise(
    `Total cases` = max(cum_cases),
    `Total deaths` = max(cum_deaths),
    Country = country
  ) %>%
  select(Country, `Total cases`, `Total deaths`) %>%
  arrange(desc(`Total deaths`))
bd_sa_summary %>%
  kable(format = "html", caption=paste0("Bangladesh and its peers: total cases and deaths compared on day ", bd_day_indicator)) %>%
  kable_styling(position = "center") %>%
  row_spec(which(bd_sa_summary$Country %in% c('Bangladesh', 'Bangladesh(unoff)')), bold = T, color = "white", background = "#f49a86")
```
