
# Bangladesh {#bangladesh}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("covid19_bd.R")
```

This iniitial report is created based on a publicly available data. The data are available on a Google Sheet. Please see the Data Source section for link. 


## Bangladesh official data (TBD)

Please see the graph below.

## Report based on publicly available data

Please note that publicly available data are __not official__ and __MAY BE INACCURATE. I personally did not verify them. Use at your discretion.


```{r, fig.width=8, fig.height=7, echo = FALSE, warning=FALSE}
a <- df_ %>% filter(varname != "cum_tests")

# Plot
p <- ggplot(a, aes(x = date, y = cum_cases, color=varname)) +
  geom_line(size=1.1) +
  theme_minimal(base_size = 14) +
  # theme(legend.position = "none") +
  # theme_void() +
  # theme(legend.position="none") +
  ggtitle("Bangladesh: COVID19 cumulative cases\n(SUSPECTED cases are UNOFFICIAL)",
          data_last_refreshed ) +
  xlab("Date") +
  ylab("Cumulative cases") +
  theme(legend.position="bottom") +
  labs(caption = "SOURCE: https://cutt.ly/StTIxJQ")

p + scale_color_discrete(name="Type",
                       breaks=c("cum_deaths_suspected", 
                                "cum_deaths_govt",
                                "cum_change_deaths_susp",
                                "cum_change_deaths_govt"),
                       
                       labels=c("Deaths (suspected)", 
                                "Deaths (confirmed)", 
                                "Daily change (suspected)",
                                "Daily change (confirmed)")
) +
  guides(fill=guides(colour=guide_legend(ncol=2,nrow=2,byrow=TRUE)))

```

```{r, echo=FALSE, message=FALSE}
# Bangladesh summary table

df %>%
  arrange(date) %>%
  summarise(
    `Total Deaths (Official)` = max(cum_deaths_govt)
    , `Total Deaths (Suspected)` = max(cum_deaths_suspected)
    , `Total Tests` = max(total_tests, na.rm = TRUE)
  ) %>% kable(caption=paste("Bangladesh summary as of ", data_last_refreshed))

```
