
# Projections {#projection}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# source("covid19_bd.R")
# source("covid19_global.R")
```


## Projection for Bangladesh (unofficial) {#projection-bangladesh}

__Last updated:__ `r today()`

This is a dynamic document and will be updated frequently. Please check back for latest projection. This projection is based on unofficial Bangladesh incidence data.

### Methodology

This projection is based on an SEIR model. Here, the **S** stands for Susceptible, **E** stands for Exposed/infected but asymptomatic, **I** stands for Infected **and** symptomatic, and **R** stands for Recovered. N is the population size.

Assuming there is no births or deaths in a population, (known as a closed population), the model is formulated by the following differential equations.


$$
\begin{align}
\frac{\partial S}{\partial t} & = -\frac{\beta I S}{N} \\
\frac{\partial E}{\partial t} &= \frac{\beta I S}{N} -\sigma E \\
\frac{\partial I}{\partial t} &= \sigma E - \gamma I \\
\frac{\partial R}{\partial t} &= \gamma I
\end{align}
$$

Here the parameters $\beta$ controls the transition from susceptible to the Exposed compartment, $\sigma$ controls the transition from E to I, and $\gamma$ controls the transition from I to R. 

The reproduction rate, $R_0$ can be approximated by 
$$R_0 = \frac{\beta}{\gamma}$$

Model was fitted using all but the last two day's incidences to obtain the estimated $\beta$ and $\gamma$. The fitted model was used for prediction. This post was inspired by @churches2020covid19rpart1, where you will find some details on the computation.


### Ascertainment Rate

Not all the cases are reported or tested. Usually a fraction of the actual cases are detected. This is known as ascertainment rate. We consider 25\%, 50\%, 75\% amd 90\% ascertainment rate when fitting the model. 

Simply, the incidences are inflated by the inverse of the ascertainment rate.



```{r, projection-bangladesh-fn,  echo = FALSE, warning=FALSE}

library(deSolve)
library(grid)
library(gridExtra)


######################################
## SIER Modeling -------
######################################
# Parameters
# beta = rate of expusore from susceptible infected contact 
# sigma = rate at which exposed person becomes infected
# gamma = rate at which infected person recovers
# S = Initial susceptible population
# E = Initial exposed population
# I = Initial infected population
# R = Recovered population

SEIR <- function(time, state, parameters) {
  par <- as.list(c(state, parameters))
  with(par, {
    dS <- -beta * I * S/N
    dE <- beta * I * S/N - sigma * E
    dI <- sigma * E - gamma * I
    dR <- gamma * I
    list(c(dS, dE, dI, dR))
  })
}


# define a function to calculate the residual sum of squares
# (RSS), passing in parameters beta and gamma that are to be
# optimised for the best fit to the incidence data

RSS <- function(parameters) {
  names(parameters) <- c("beta", "sigma", "gamma")
  out <- ode(y = init, times = Day, func = SEIR, parms = parameters)
  fit <- out[, 4]
  sum((infected - fit)^2)
}


# put the daily cumulative incidence numbers for USA from
# 15th Jan to 30th Jan into a vector called Infected
# 

fit_seir <- function(country_name='Bangladesh(unoff)', N=170000000, af=0.5, npast=2, nfuture=10){
  # country = Country name
  # N = population size of the country
  # af = ascertainment factor, default = 0.5
  # country = "Bangladesh(unoff)"
  # npast = number of days in the past to exclude when fitting the model
  # default is npast = 5
  # nfuture = number of days in the future the algorithm to predict to
  # default is nfuture=10
  
  
  SEIR <- function(time, state, parameters) {
    par <- as.list(c(state, parameters))
    with(par, {
      dS <- -beta * I * S/N
      dE <- beta * I * S/N - sigma * E
      dI <- sigma * E - gamma * I
      dR <- gamma * I
      list(c(dS, dE, dI, dR))
    })
  }
  
  
  # define a function to calculate the residual sum of squares
  # (RSS), passing in parameters beta and gamma that are to be
  # optimised for the best fit to the incidence data
  
  RSS <- function(parameters) {
    names(parameters) <- c("beta", "sigma", "gamma")
    out <- ode(y = init, times = Day, func = SEIR, parms = parameters)
    fit <- out[, 4]
    sum((infected - fit)^2)
  }
  
  
  country = enquo(country_name)
  df <- bd_unoff %>% filter(country == !!country, cum_cases>0)
  infected <- df %>% filter(date >= min(date), date <= today() - 1 - npast) %>% 
    pull(cum_cases)
  
  R = 0; E=0; I = infected[1]; S = N - E - I - R
  
  seir_start_date <- df %>% pull(date) %>% min()
  
  # Ascertainment factor
  infected = infected * 1/af

  
  # Create an incrementing Day vector the same length as our
  # cases vector
  Day <- 1:(length(infected))
  
  # now specify initial values for S, I and R
  init <- c(S = S, E=E, I=I, R=R)
  
  # now find the values of beta and gamma that give the
  # smallest RSS, which represents the best fit to the data.
  # Start with values of 0.5 for each, and constrain them to
  # the interval 0 to 1.0
  
  opt <- optim(c(.5, .5, .5), RSS, method = "L-BFGS-B", 
               lower = c(0.01,0.01,0.01), upper = c(.999, .999, .999), 
               control=list(maxit = 1000))
  
  # check for convergence
  opt_msg = opt$message
  opt_par <- setNames(opt$par, c("beta", "sigma", "gamma"))

  beta = opt_par["beta"]
  gamma = opt_par["gamma"]
  sigma = opt_par["sigma"]
  R0 = as.numeric(beta/gamma)
  
  
  # time in days for predictions
  t <- 1:(as.integer(today() - seir_start_date)  + nfuture)
  
  # get the fitted values from our SEIR model
  
  odefit = ode(y = init, times = t, func = SEIR, parms = opt_par)
  fitted_cases <- data.frame(odefit)
  
  # add a Date column and join the observed incidence data
  fitted_cases <- fitted_cases %>% 
    mutate(date = seir_start_date + days(t - 1)) %>% 
    left_join(df %>% filter(cum_cases>0) %>% ungroup() %>%
                select(date, cum_cases))
  
  
  # Return
  list(country=country_name, infected = infected,
       opt_msg=opt_msg, opt_par = opt_par, R0=R0, opt_msg=opt_msg, 
       fitted_cases=fitted_cases, N=N, af=af)
  
}

```


```{r, projection-bangladesh-graph, fig.width=10, fig.height=8, echo = FALSE, warning=FALSE, cache=TRUE, message=FALSE}

a1 <- fit_seir(country = "Bangladesh(unoff)",  af=.25, nfuture = 5)
p1<- a1$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a1$R0, digit=3),"and","assumsed reporting =", a1$af*100,"%" ))

a2 <- fit_seir(country = "Bangladesh(unoff)", af=.5, nfuture = 5)
p2<- a2$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a2$R0, digit=3),"and","assumsed reporting =", a2$af*100,"%" ))

a3 <- fit_seir(country = "Bangladesh(unoff)", af=.75, nfuture = 5)
p3<- a3$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a3$R0, digit=3),"and","assumsed reporting =", a3$af*100,"%" ))


a4 <- fit_seir(country = "Bangladesh(unoff)", af=.9, nfuture = 5)
p4<- a4$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a4$R0, digit=3),"and","assumsed reporting =", a4$af*100,"%" ))



grid.arrange(grobs = list(p1, p2, p3, p4), ncol=2, 
             top = textGrob(paste0("COVID-19 observed and 5-day predicted cumulative incidence for Bangladesh (unofficial)\n(orange=actual cases, red=predicted incidence from SEIR model, blue=Exposed)\n",  today()),
             gp=gpar(fontsize=16, font=16)))
```

It turns out that the bottom left one fits the current data best. So lets put that figure on a bigger canvas.


```{r, projection-bangladesh-graph-best, fig.width=10, fig.height=8, echo = FALSE, warning=FALSE}
# Plotting the best fit
p3
```

### Projections for the next 5 days

```{r, echo=FALSE, fig.cap="Actual and Predicted Cumulative Cases using SEIR Model"}
a3$fitted_cases %>% as.data.frame() %>%
  mutate(
    Date = date,
    `Actual daily cases` = c(0, diff(cum_cases)),
    `Actual cumulative cases` = cum_cases,
    `Projected cumulative cases` = round(I, digits = 0),
    `Projected daily cases` = c(0, diff(`Projected cumulative cases`))
  ) %>%
  select(Date, `Actual daily cases`, `Projected daily cases`,
         `Actual cumulative cases`,
          `Projected cumulative cases` ) %>%
  tail(10) %>%
  kable(format = "html", 
        caption=paste("Predicted new cases for the next 5 days")) %>%
          kable_styling(position = "center")
```

### Projection for 100 days into the future

Assuming the situation will remain like this including the interventions currently in place, the 100 day projection suggests that the the peak of the epidemic will be around the middle of June. The trajectory also suggests that the epidemic will end by end of July or early August. 

```{r, projection-bangladesh-graph-100, fig.width=10, fig.height=8, echo = FALSE, warning=FALSE, cache=TRUE, message=FALSE}

nfuture=100

a1 <- fit_seir(country = "Bangladesh(unoff)", af=.25, nfuture = nfuture)
p1<- a1$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  # geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a1$R0, digit=3),"and","assumsed reporting =", a1$af*100,"%" ))

a2 <- fit_seir(country = "Bangladesh(unoff)", af=.5, nfuture = nfuture)
p2<- a2$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  # geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a2$R0, digit=3),"and","assumsed reporting =", a2$af*100,"%" ))

a3 <- fit_seir(country = "Bangladesh(unoff)", af=.75, nfuture = nfuture)
p3<- a3$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  # geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a3$R0, digit=3),"and","assumsed reporting =", a3$af*100,"%" ))


a4 <- fit_seir(country = "Bangladesh(unoff)", af=.9, nfuture = nfuture)
p4<- a4$fitted_cases %>% 
  ggplot(aes(x = date)) + 
  geom_line(aes(y = I), colour = "red") +
  geom_line(aes(y = E), colour = "blue") +
  # geom_line(aes(y = R), colour = "darkgreen") +
  geom_point(aes(y = cum_cases), colour = "orange") + 
  scale_y_continuous(labels = scales::comma) +
  # scale_y_continuous(labels = scales::comma, trans = 'log') + 
  labs(y = "Cumulative incidence", 
       title = paste("R0 =", format(a4$R0, digit=3),"and","assumsed reporting =", a4$af*100,"%" ))


grid.arrange(grobs = list(p1, p2, p3, p4), ncol=2, 
             top = textGrob(paste0("COVID-19 observed and 100-day predicted cumulative incidence for Bangladesh (unofficial)\n(orange=actual cases, red=predicted incidence from SEIR model, blue=Exposed)\n",  today()),
             gp=gpar(fontsize=16, font=16)))

# grid.arrange(a, b, c, d,ncol=2, 
#      nrow=2, top=textGrob("Title", gp=gpar(fontsize=15,font=8)))

```

### R code for fitting SEIR

Please let me know if you find any error in it. The code was adapted from @churches2020covid19rpart1

```{r,  echo = TRUE, warning=FALSE, eval=FALSE}

library(deSolve)
library(grid)
library(gridExtra)


######################################
## SIER Modeling -------
######################################
# Parameters
# beta = rate of expusore from susceptible infected contact 
# sigma = rate at which exposed person becomes infected
# gamma = rate at which infected person recovers
# S = Initial susceptible population
# E = Initial exposed population
# I = Initial infected population
# R = Recovered population

SEIR <- function(time, state, parameters) {
  par <- as.list(c(state, parameters))
  with(par, {
    dS <- -beta * I * S/N
    dE <- beta * I * S/N - sigma * E
    dI <- sigma * E - gamma * I
    dR <- gamma * I
    list(c(dS, dE, dI, dR))
  })
}


# define a function to calculate the residual sum of squares
# (RSS), passing in parameters beta and gamma that are to be
# optimised for the best fit to the incidence data

RSS <- function(parameters) {
  names(parameters) <- c("beta", "sigma", "gamma")
  out <- ode(y = init, times = Day, func = SEIR, parms = parameters)
  fit <- out[, 4]
  sum((infected - fit)^2)
}


fit_seir <- function(country_name='Bangladesh(unoff)', 
                     N=170000000, af=0.5, npast=2, nfuture=10){
  # country = Country name
  # N = population size of the country
  # af = ascertainment factor, default = 0.5
  # country = "Bangladesh(unoff)"
  # npast = number of days in the past to exclude when fitting the model
  # default is npast = 5
  # nfuture = number of days in the future the algorithm to predict to
  # default is nfuture=10
  
  
  SEIR <- function(time, state, parameters) {
    par <- as.list(c(state, parameters))
    with(par, {
      dS <- -beta * I * S/N
      dE <- beta * I * S/N - sigma * E
      dI <- sigma * E - gamma * I
      dR <- gamma * I
      list(c(dS, dE, dI, dR))
    })
  }
  
  # define a function to calculate the residual sum of squares
  # (RSS), passing in parameters beta, sigma, and gamma that are to be
  # optimised for the best fit to the incidence data
  
  RSS <- function(parameters) {
    names(parameters) <- c("beta", "sigma", "gamma")
    out <- ode(y = init, times = Day, func = SEIR, parms = parameters)
    fit <- out[, 4]
    sum((infected - fit)^2)
  }
  
  country = enquo(country_name)
  df <- bd_unoff %>% filter(country == !!country, cum_cases>0)
  infected <- df %>% filter(date >= min(date), date <= today() - 1 - npast) %>%
    pull(cum_cases)
  
  R = 0; E=0; I = infected[1]; S = N - E - I - R
  
  seir_start_date <- df %>% pull(date) %>% min()
  
  # Ascertainment factor
  infected = infected * 1/af

  # Create an incrementing Day vector the same length as our
  # cases vector
  Day <- 1:(length(infected))
  
  # now specify initial values for S, I and R
  init <- c(S = S, E=E, I=I, R=R)
  
  # now find the values of beta, sigma. and gamma that give the
  # smallest RSS, which represents the best fit to the data.
  # We perform a constrined optimization with initial parameters
  # set as .5. We bound the params within 0.01 and 0.999 (I've made 
  # the choice). Feel free to change it to c(0, 1). But you may face
  # issues with convergence
  
  opt <- optim(c(.5, .5, .5), RSS, method = "L-BFGS-B", 
               lower = c(0.01,0.01,0.01), upper = c(.999, .999, .999), 
               control=list(maxit = 1000))
  
  # check for convergence
  opt_msg = opt$message
  opt_par <- setNames(opt$par, c("beta", "sigma", "gamma"))

  beta = opt_par["beta"]
  gamma = opt_par["gamma"]
  sigma = opt_par["sigma"]
  R0 = as.numeric(beta/gamma) # there is a more complex formula 
  
  
  # time in days for predictions
  t <- 1:(as.integer(today() - seir_start_date)  + nfuture)
  
  # get the fitted values from our SEIR model
  
  odefit = ode(y = init, times = t, func = SEIR, parms = opt_par)
  fitted_cases <- data.frame(odefit)
  
  # add a Date column and join the observed incidence data
  fitted_cases <- fitted_cases %>% 
    mutate(date = seir_start_date + days(t - 1)) %>% 
    left_join(df %>% filter(cum_cases>0) %>% ungroup() %>%
                select(date, cum_cases))
  
  
  # Return
  list(country=country_name, infected = infected,
       opt_msg=opt_msg, opt_par = opt_par, R0=R0, opt_msg=opt_msg, 
       fitted_cases=fitted_cases, N=N, af=af)
  
}
```

